Scheduling and Vectorization for MPC

### 2022-02-21 paper_SIMD.tex
Minor edits.

%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title[Scheduling and Amortization for MPC]{Scheduling and Amortization for MPC}

@@

%
% Keywords. The author(s) should pick words that accurately describe the work being
% presented. Separate the keywords with commas.
\keywords{protocol mixing; linear programming; multiparty computation; program analysis; cryptography}

@@

As a running example, we assume a standard MPC benchmark, Biometric matching.
An intuitive implementation is as follows:

{\small
\begin{verbatim}
def biometric(C: shared[list[int]], D: int, 
              S: shared[list[int]], N: int) -> 
                      tuple[shared[int], shared[int]]:
    min_sum = 10000
    min_index = 0  
    for i in range(N):
        sum = 0
        for j in range(D):
            d = S[i * D + j] - C[j]
            p = d * d
            sum = sum + p
        if sum < min_sum:
            min_sum = sum
            min_index = i
    return (min_sum, min_index)
\end{verbatim}
}

Array \texttt{C} is the feature vector of \texttt{D} 
features that we wish to match and array \texttt{S} is the database of \texttt{N} vectors of size \texttt{D}
that we match against. The programmer annotates the inputs and outputs as \emph{shared}.
The program iterates over the entries in the database and computes
the sum of squares of the differences of individual features. The program returns the 
index \texttt{i} of the vector that gives the best match plus the corresponding sum of squares.

Our compiler imposes the following restrictions. We note that in most cases, the restrictions 
can be easily lifted and we plan to do so in future iterations of our compiler.

\begin{enumerate}
\item The program contains arbitrarily nested loops, however, loop bounds are fixed: \texttt{0 <= i < N}.
A standard restriction in MPC is that the bounds must be known at circuit-generation time.
\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
in row-major order and at this point the programmer is responsible for linearization
and access. 
\item Array subscrpts are plaintext values.
\item Our compiler allows for output (write) arrays, however it restricts
write access to \emph{canonical writes} along the dimensions of the array (i.e., \texttt{A[i,j] = ...}
where \texttt{i} and \texttt{j} loop over the two dimensions of \texttt{A} is allowed, 
but \texttt{A[i,j+2] = ...} is not allowed). Read access is arbitrary. 
\end{enumerate}

@@

We linearize the source turning conditionals into MUX statements. The PHI nodes
are remnants of the SSA IR; our compiler generates code that picks the correct value 
when producing MOTION output and the MOTION framework in turn linearizes 
MPC source loops when it generates the circuit.

@@

Our key contribution is the vectorizing transformation. We can compute all $N*D$ 
subtraction operations (line 9) in a single SIMD instruction; similarly we can compute 
all multiplication operations (line 10) in a single SIMD instruction. And while we cannot
vectorize computation of the N individual sums, we can compute the N sums in parallel. 
Our compiler automatically detects these opportunities and transforms the program. 
It is standard that MPC researchers write vectorized versions of the Biometric
program by hand; we are the first (to the best of our knowledge) to automatically 
transform an intuitive, iterative MPC program into an unintuitive vectorized one.

\subsection{Vectorization and Cost of Schedule}

Our compiler produces the following vectorized program. (Note that this is
still higher-level IR, our compiler turns this code into MOTION variables, 
loops and SIMD primitives, which MOTION then uses to generate the circuit.)

>>>>

%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title[Scheduling and Vectorization for MPC]{Scheduling and Vectorization for MPC}

@@

%
% Keywords. The author(s) should pick words that accurately describe the work being
% presented. Separate the keywords with commas.
\keywords{multiparty computation; compilers; cryptography}

@@

As a running example, consider Biometric matching, a standard MPC benchmark.
Array \texttt{C} is the feature vector of \texttt{D} features that we wish to match and array \texttt{S} 
is the database of \texttt{N} vectors of size \texttt{D} that we match against.
An intuitive implementation is as follows:

{\small
\begin{verbatim}
def biometric(C: shared[list[int]], D: int, 
              S: shared[list[int]], N: int) -> 
                      tuple[shared[int], shared[int]]:
    min_sum = 10000
    min_index = 0  
    for i in range(N): #loop over database
        sum = 0
        for j in range(D): #loop over features
            d = S[i * D + j] - C[j] #i.e., d = S[i,j] - C[j]
            p = d * d
            sum = sum + p
        if sum < min_sum:
            min_sum = sum
            min_index = i
    return (min_sum, min_index)
\end{verbatim}
}

Our compiler takes (essentially) standard IMP syntax.
The programmer can write intuitive iterative programs as the one above.
They annotate certain inputs and outputs as \emph{shared}.
Here the code iterates over the entries in the database and computes
the sum of squares of the differences of individual features. The program returns the 
index \texttt{i} of the vector that gives the best match plus the corresponding sum of squares.

Our compiler imposes the following restrictions. We note that in some cases, the restrictions 
can be easily lifted and we plan to do so in future iterations of our compiler.

\begin{enumerate}
\item The program contains arbitrarily nested loops, however, loop bounds are fixed: \texttt{0 <= i < N}.
A standard restriction in MPC is that the bounds must be known at circuit-generation time.
\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
in row-major order and at this point the programmer is responsible for linearization
and access. 
\item Array subscrpts are plaintext values.
\item Our compiler allows for output (write) arrays, however it restricts
write access to \emph{canonical writes} along the dimensions of the array. I.e., \texttt{A[i,j] = ...}
where \texttt{i} and \texttt{j} loop over the two dimensions of \texttt{A} is allowed, 
but \texttt{A[i,j+2] = ...} is not allowed. Read access is arbitrary. 
\end{enumerate}

@@

The compiler linearizes the source turning conditionals into MUX statements. The PHI nodes
are remnants of the SSA IR; the compiler generates code that picks the correct value 
when producing MOTION output; the MOTION framework in turn linearizes 
loops when it generates the circuit.

@@

A key contribution is the vectorizing transformation. We can compute all $N*D$ 
subtraction operations (line 9) in a single SIMD instruction; similarly we can compute 
all multiplication operations (line 10) in a single SIMD instruction. And while we cannot
vectorize computation of the N individual sums, we can compute the N sums in parallel. 
Our compiler \emph{automatically detects these opportunities and transforms the program}. 
It is standard that MPC researchers write vectorized versions of the Biometric
program by hand; we are the first (to the best of our knowledge) to automatically 
transform an intuitive, iterative MPC program into an unintuitive vectorized one.

\subsection{Vectorized MPC Source and Cost of Schedule}

Our compiler produces the following vectorized program. (Note that this is
still higher-level IR, Vectorized MPC Source. Our compiler turns this code into MOTION variables, 
loops and SIMD primitives, which MOTION then uses to generate the circuit.)
